---
title: "Report of index aggregation"
author: "Luna Zhang, AbacusBio"
date: "`r Sys.Date()`"
# always_allow_html: true
output:
  html_document: 
    toc: true
    toc_depth: 3
    toc_float: true
  word_document:
    toc: true # 
    reference_docx: template_ab.docx # template_default_tableStyle.docx #
  # pdf_document:
    # toc: true
toc-title: Table of Contents # https://community.rstudio.com/t/rmarkdown-into-word-change-a-title-in-toc/48137/4
params:
  demo: 0
  file_name_ebv: NA
  file_name_ev: NA
  file_name_wt: NA
  filter_col_ebv: NA
  filter_level_ebv: NA # reactive
  filter_col_ev: NA
  filter_level_ev: NA # rctv
  filter_ebv_na_0: FALSE
  sumstat_ebv_table: NA
  sumstat_ebv_digit: 2
  sumstat_ebv_p: NA
  sumstat_ebv_table_str: NA
  sumstat_ebv_p_str: NA
  sumstat_ev_table: NA
  sumstat_ev_digit: 2
  sumstat_ev_p: NA
  sumstat_ev_table_str: NA
  sumstat_ev_p_str: NA
  n_indi: NA
  n_index: NA
  sumstat_index: NA
  sumstat_index_digit: 2
  sumstat_index_p: NA
  cl_input: NA
  cl_k: 2
  cl_best_method: NA
  cl_agg_coefs: NA
  cl_agg: NA
  cl_op_cut: NA # rctv? $p_tss, p$sil
  cl_dendro: NA
  cl_cor: NA
  cl_cor_digit: NA
  cl_cor_p: NA
  cl_show_cor: TRUE
  cl_heat: NA
  # cl_heat_data: NA
  cl_choose_wt: NA
  new_index_wt: NA
  new_ew_digit: NA
  new_ew: NA
  new_ew_p: NA
  new_relew: NA
  new_relew_p: NA
  dx_bench: NA
  dx_cor: NA
  dx_cor_p: NA
  dx_sel_agg: NA
  dx_cor_title: NA
  dx_cor_selected: NA
  dx_cor_p_selected: NA
  dx2_top_n_title: NA
  dx2_top_n_table: NA
  dx2_top_n_p: NA
  dx3_classvar: NA
  dx3_agg_by: NA
  dx3_summary: NA
  dx3_summary_p: NA
  dx4_wtvar: NA
  dx4_agg_by: NA
  dx4_summary: NA
  dx4_summary_p: NA
---

```{r setup, include=FALSE}
# how to change word doc formats:
## https://rmarkdown.rstudio.com/articles_docx.html

# kableExtra styling:
# https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html#Group_rows_via_labeling
options(kableExtra.auto_format = FALSE) # ensure tables in word doc show normally
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE) # , root.dir = "R/")

library(dplyr)
library(ggplot2)
# library(kable)
library(kableExtra)
```

# Disclaimer

This report is generated from the **Plant Index Aggregation App**. You may notice some `NULL` or 
`NA` in the report. It means the corresponding analysis was skipped when the app was used. 

# Background

The objective of this analysis is to create aggregated indexes that can represent clusters of 
numeric original indexes. The report contains the information of the original and aggregated 
indexes, the process of cluster forming, and the evaluation of the performance of the aggregated 
indexes.
  
# Input 

## Files

```{r input}
# print(getwd()) # "C:/Users/lzhang/AppData/Local/Temp/RtmpOuA4e0" == tempDir()
# print(knitr::opts_chunk$get("root.dir")) # NULL
files <- c(params$file_name_ebv, params$file_name_ev, params$file_name_wt)

if (params$demo > 0) {
  files <- rep("demo", 3)
}

data.frame(Table = c("Breeding value
                     ", "Economic value", "Economic value weight"),
           File = files) %>% 
  knitr::kable() %>% # kableExtra::kbl() doesn't show table format in word doc
  kable_minimal(full_width = F, position = "left")
```

# Filters

Missing EBV set to 0:, `r params$filter_ebv_na_0`

```{r filter}

if(!is.null(params$filter_col_ebv)) {
  filter_col_ebv <- params$filter_col_ebv
  filter_level_ebv <- params$filter_level_ebv()
} else {
  filter_col_ebv <- ""
  filter_level_ebv <- "" # error if use params$filter_level_() directly when it's not shown on UI
}

if(!is.null(params$filter_col_ev)) {
  filter_col_ev <- params$filter_col_ev
  filter_level_ev <- params$filter_level_ev()
} else {
  filter_col_ev <- ""
  filter_level_ev <- ""
}

tibble::tibble(Table = c(rep("Breeding value",
                             max(1, length(params$filter_col_ebv))),
                         rep("Economic value",
                             max(1, length(params$filter_col_ev)))),
               column = c(filter_col_ebv, filter_col_ev),
               level = unlist(list(filter_level_ebv, filter_level_ev), recursive = F)
) %>%
  knitr::kable(caption = "Table filters") %>% 
  kable_minimal(full_width = F, position = "left")
```

# Summary statistics

## Breeding value

```{r sumstat_ebv}
params$sumstat_ebv_table %>% 
  knitr::kable(caption = "Breeding value summary statistics", digits = params$sumstat_ebv_digit) %>% 
  kable_minimal()

# height <- ifelse(length(params$sumstat_ebv_p) > 4, 12, 6)
```

```{r sumstat_ebv_p, warning=TRUE}

if(class(params$sumstat_ebv_p)[1]=="list") {
  gridExtra::grid.arrange(grobs = params$sumstat_ebv_p,
                          ncol = min(4, length(params$sumstat_ebv_p))) %>%
    invisible()
} else {
    params$sumstat_ebv_p
  }
```


### Classification variable {.tabset}

#### Distribution

```{r sumstat_ebv_p_str}

if(class(params$sumstat_ebv_p_str)[1]=="list") {
  gridExtra::grid.arrange(grobs = params$sumstat_ebv_p_str,
                          ncol = min(4, length(params$sumstat_ebv_p_str))) %>% 
    invisible()
} else {
    params$sumstat_ebv_p_str 
  }
```

#### Full table

```{r sumstat_ebv_str}
params$sumstat_ebv_table_str %>% 
  knitr::kable(caption = "classification variables summary statistics", 
      digits = params$sumstat_ebv_digit) %>%
  kable_minimal()
```


## Economic value

```{r sumstat_ev}
params$sumstat_ev_table %>% 
  knitr::kable(caption = "Economic value summary statistics", digits = params$sumstat_ebv_digit) %>% 
  kable_minimal()

# height <- ifelse(length(params$sumstat_ebv_p) > 4, 12, 6)
```

```{r sumstat_ev_p, warning=TRUE}

if(class(params$sumstat_ev_p)[1]=="list") {
  gridExtra::grid.arrange(grobs = params$sumstat_ev_p,
                          ncol = min(4, length(params$sumstat_ev_p))) %>%
    invisible()
} else {
    params$sumstat_ev_p
  }
```


### Classification variable {.tabset}

#### Distribution

```{r sumstat_ev_p_str}

if(class(params$sumstat_ev_p_str)[1]=="list") {
  gridExtra::grid.arrange(grobs = params$sumstat_ev_p_str,
                          ncol = min(4, length(params$sumstat_ev_p_str))) %>% 
    invisible()
} else {
    params$sumstat_ev_p_str 
  }
```

#### Full table

```{r sumstat_ev_str}
params$sumstat_ev_table_str %>% 
  knitr::kable(caption = "classification variables summary statistics", 
      digits = params$sumstat_ev_digit) %>%
  kable_minimal()
```


## Original indexes {.tabset}

You have `r params$n_indi` individuals and `r params$n_index` original indexes.  
Individuals with missing breeding value are removed.
           
### Distribution of 100 random samples

```{r sumstat_index_p, fig.dim=c(12,6)}
params$sumstat_index_p
```

### Full table

```{r sumstat_index}
params$sumstat_index %>% 
  knitr::kable(caption = "Original index summary statistics", 
      digits = params$sumstat_index_digit) %>%
  kable_minimal()
```

# Clustering analysis

## Clusters {.tabset}

The *`r params$cl_input`* matrix is used as the input. Hierarchical clustering with `r params$agg`  agglomeration algorithm is used to compute the clustering. $k = `r params$cl_k`$ clusters was chosen in the end.

### Result

```{r cl, fig.dim=c(12,6)}
params$cl_dendro
```

The full table is downloaded separately.

### Parameter Tuning

The best agglomeration method is `r params$cl_best_method` with the largest agglomerative 
coefficient of among other methods:

```{r cl_par}
print(params$cl_agg_coefs)
```

There are three ways to determine the best k:

#### by merging steps:

```{r cl_par0}
# if par tuning isn't run, params$cl_op_cut() is NULL instead of a reactive function
if(is.null(params$cl_op_cut)) {
  kh = h = p_tss = p_sil <- NULL
} else {
  kh <- params$cl_op_cut()$k_h
  h <- params$cl_op_cut()$h
  p_tss <- params$cl_op_cut()$p_tss
  p_sil <- params$cl_op_cut()$p_sil
}
```

The best k is `r kh` at the largest height change of `r h`.

#### by largest total within-cluster sum of squares drop:

```{r cl_par1}
p_tss
```

#### by largest mean within-cluster silhouette value

```{r cl_par2}
p_sil
```


## Summary statistics

### Within Cluster correlation distribution

```{r cl_sumstat}
params$cl_cor %>%
  knitr::kable(caption = "Within Cluster correlation",
      digits = params$cl_cor_digit) %>%
  kable_minimal()
  
```

```{r cl_sumstat_p}
params$cl_cor_p
```

## Diagnosis

```{r cl_diagnosis}
txt <- NA

if(!is.null(params$cl_show_cor)) {
  if(params$cl_show_cor) {
    txt <- "index correlation"
    } else {
      txt <- "individual by index"
    }
  }
```

Below is the *`r txt`* matrix heatmap.

```{r cl_diagnosis1}
params$cl_heat
```

The full correlation matrix can be downloaded separately.

# Aggregated index

## Index weight {.tabset}

The *`r params$cl_choose_wt`* was used to weigh original indexes within their cluster.

### Sneak peek 

```{r new_wt}
if(!is.null(params$new_index_wt)) {
 params$new_index_wt %>%
  head() %>%
  knitr::kable() %>%
  kable_minimal(full_width = F, position = "left")
}
```

### full table

```{r new_wt1}
params$new_index_wt %>%
  knitr::kable() %>%
  kable_minimal(full_width = F, position = "left")
```


## Aggregated index economic values

```{r new_ev}
params$new_ew %>%
  knitr::kable(digits = params$new_ew_digit, caption = "New economic values") %>%
  kable_minimal(full_width = F, position = "left")
```

```{r new_ev_p, fig.dim=c(12,8)}
params$new_ew_p
```

## Aggregated index relative economic weight (RW)

$RW = EV \times SD(EBV)$

```{r new_rel_ew}
params$new_relew %>%
  knitr::kable(digits = params$new_ew_digit, caption = "New relative economic weight") %>%
  kable_minimal(full_width = F, position = "left")
```

```{r new_rel_ew_p, fig.dim=c(12,8)}
params$new_relew_p
```

# Aggregated index diagnosis

This is to investigate the relationships between original and aggregated indexes, and across 
aggregated indexes, with or without comparison with a benchmark index, to evaluate how well 
aggregated indexes represent their own cluster.

## Correlation

The ideal outcome is that the majority of the correlations of original indexes with their
corresonding aggregated indexes are higher than the correlations of original indexes with a 
benchmark index.

### Distribution of correlations between within-cluster indexes and their aggregated index

#### All aggregated index and the benchmark index {.tabset}

Below is the sorted correlations between within-cluster original indexes and their aggregated 
indexes (coloured dots) vs. those between original indexes and the benchmark index: 
*`r params$dx_bench`*.

##### plot

```{r dx_corr}
params$dx_cor_p
```

##### full table

```{r dx_corr_table}
params$dx_cor %>%
  knitr::kable(digits = 3, 
      caption = "Correlation between original and aggregated indexes") %>%
  kable_minimal(full_width = F, position = "left")
```


#### Selected indexes {.tabset}

Below is the sorted `r params$dx_cor_title`

##### plot

```{r dx_corr1}
params$dx_cor_p_selected
```

##### full table

```{r dx_corr1_table}
params$dx_cor_selected %>%
  knitr::kable(digits = 3, 
      caption = "Correlation between selected original and aggregated indexes") %>%
  kable_minimal(full_width = F, position = "left")
```


## Top individual rank

The overlap of the top individuals in different indexes is an indication of how similar they are.
Ideally, the overlap should be as small as possible across different aggregated indexes. Too much 
overlap between two indexes implies they can be merged.

### `r params$dx2_top_n_title` {.tabset}

#### plot

The top individual % overlap between original and aggregated indexes.

```{r top_indi}
params$dx2_top_n_p
```

#### full table

```{r top_indi1}
params$dx2_top_n_table %>%
  knitr::kable(digits = 3, 
      caption = "The top n individual overlap between original and aggregaed indexes") %>%
  kable_minimal(full_width = F, position = "left")
```

## Variable pattern {.tabset}

The frequency of levels of a known variable in aggregated indexes. This can potentially provide 
insights into the possible reason behind the aggregation, if the frequency distribution of the 
known variable is uneven across aggregated indexes.

Below is the frequency/percent distribution of *`r params$dx3_classvar`* (variable) across 
*`r params$dx3_agg_by`* (aggregated or pre-defined grouping index).

### plot

```{r classvar_pattern, fig.height=8}
params$dx3_summary_p
```

### full table

```{r classvar_pattern1}
params$dx3_summary %>%
  knitr::kable(digits = 3, 
      caption = paste0("Distribution of ", params$dx3_classvar, " across ", params$dx3_agg_by)) %>%
  kable_minimal(full_width = F, position = "left")
```

## Index weight pattern {.tabset}

The weight of each aggregated index as the sum of the weights of its corresponding original indexes
of the same cluster. This can show which aggregated index was assigned the most or the least 
weight, which is often related to the importance of the aggregated index.

Below is the distribution of *`r params$dx4_wtvar`* (index weight) across *`r params$dx4_agg_by`*
 (aggregated or pre-defined grouping index).
 
### plot

```{r weight_pattern}
params$dx4_summary_p
```

### full table

```{r weight_pattern1}
params$dx4_summary %>%
  knitr::kable(digits = 3, 
      caption = paste0("Distribution of ", params$dx4_wtvar, " across ", params$dx4_agg_by)) %>%
  kable_minimal(full_width = F, position = "left")
```