---
title: "Preprocess"
author: "Luna Zhang"
date: "22/07/2021"
output:
  html_document:
    toc: true
    css: ../css/bootstrap_minty_edited.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
```

# Step 2: Filter EBV
This tab is used to select and filter EBV input file for the downstream analysis.

## Table in the main Tab

This is a dynamic table where you can filter directly.

Click the dropdown menu on the top left change the number of entries to show.

In the **search:** search bar on the top, type in the keyword, partial keyword or number to filter rows of the entire table.
Multiple keywords can be concatenated by `|`. e.g. `1|2`. 

Similar filtering method can be applied to each column filter individually, too, which is under the header of each *ID*, 
*ClassVar* or *Group* column.
To reverse filter, use `[^keyword]` format. For example, to remove all "M" in *sex* column, type in `[^M]`. Multiple keywords
can be concatenated together, e.g. `[^index_1^index_2]`.

The bottom left of the table shows the entries on the screen and the total number of entries.

Use the page buttons on the bottom middle to turn pages.

## Sidebar
The **Filter and save** button has to be hit to update and save the table for downstream analysis. Without hitting it, the app
won't do anything after this step. Click it whenever you finish selecting and filtering the input table.

Click the **Download the table** button to download a .csv file to your local drive.

Click the **Download the filter records** button to download your operation records. This shows what filter operation you have
done to the UI before clicking the **Filter and save** button.

### Variable display controler
This sidebar has filtering functions that the table in the main tab itself doesn't have or the operation of which is too complicated.

In **Select variables to show:** selection menu, use mouse, Shift, Ctrl, Ctrl+A etc. to select columns.

Use **Decimals** numeric input box to change the number of decimal places to show in the table. However, the full decimal places will
always be saved for the downstream analysis.

Use **Rename new data as"** input box to change the file name to be downloaded to your local drive.

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
