---
title: "Preprocess Files"
author: "Luna Zhang"
date: "21/07/2021"
# output: html_document
output:
  html_document:
    toc: true
    css: ../css/bootstrap_minty_edited.css
    code_folding: hide
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = F, eval = T)
```

This menu is for you to upload, filter and inspect your input data.

# Step 1: Upload

## Demo
First time user can click the **Run demo data** button and visit all the tabs from left to right to see the full function of the app.
The Demo data uses a corn dataset. The traits EBV and EV are simulated using mean, SD and correlaiton from a small real dataset.

## Upload data

Upload your data from the sidebar menu on the left. The files are:

- **Estimated breeding value (EBV) description file**
- **EBV file**
- **Economic value/weight (EV) description file**
- **EV file**
- **optional: index weight file**

The two description files have two columns:
*column_labelling* and *classifier*. Column_labelling column contains the column names in the input file the description file pointing to, 
e.g. EBV or EV input file. And classifier column describes the type of the corresponding column in the input file. 

Common legal classifiers include:

- **ID**: the unique ID for each row of individual (in EBV file) or index (in EV and index weight file)
- **ClassVar**: classification variable, such as a character string variable
- **EBV**: trait EBV
- **EV**: trait EV

Below is a sample EBV description file used in demo data:

```{r inputfile, warning = F}
bv_desc <- readxl::read_xlsx("../data/description_bv.xlsx")
# knitr::kable(ev_desc)
bv <- read.csv("../data/bv.csv", header = T, nrows = 6, fileEncoding = "UTF-8-BOM")
ev_desc <- read.csv("../data/description_ev.csv", header = T, fileEncoding = "UTF-8-BOM")
ev <- read.csv("../data/ev.csv", header = T, nrows = 6, fileEncoding = "UTF-8-BOM")
weight <- read.csv("../data/index_weight.csv", header = T, nrows = 6, fileEncoding = "UTF-8-BOM")
knitr::kable(bv_desc, caption = "EBV description file")
```

Optional classifiers include

- **Group**: a character string variable of user-defined group. Can be potentially used to form aggregated index

Below is a part of a sample EV description file from demo data. The *sector* variable is a *Group* classifier:

```{r EV_description}
knitr::kable(ev_desc, caption = "EV description file")
```

Below are parts of sample EBV and EV files from demo data:

```{r EV_EBV}
knitr::kable(bv, caption = "EBV file")
knitr::kable(ev, caption = "EV file")
```

Index weighting file contains the weightings for each index to be used when some of them are aggregated together:

```{r weight}
knitr::kable(weight, caption = "Index weighting file")
```

All files can be in one of `.csv/.xlsx/.txt` format, with comma or space as the delimitor.

Missing values can be empty cell or `NA`. Individuals with missing values in the EBV file will be excluded from calculating index value
unless they are imputated in the next Tab (filter). Traits with missing EV values in the EV file will be imputed as `0` by default.

## Info, warning and erro messages

In the sidebar and main tab, there will be text message prompting during demo or file uploading. For example, if the uploaded file has
illegal classifiers, columns names or other things, the warning message will show the illegal element and instruct the user to either
refresh the browswer and start over or upload a correct file.